---
title: "Lecture 6: R Functions"
author: "Sathvik Guntha (A:17927716)"
format: pdf
toc: True
---

## Background

All functions in R gave at least 3 things: 

- A **name** that we used to call the function
- One or more input **arguments**
- The **body** the lines of R code that do the work

## First Function
Let's write a silly wee function called `add()` to add some numbers (the input arguments)

```{r}
add <- function(x,y){
  x + y
}
```

```{r}
add (10, 1)
```

```{r}
add (c(10,10,10), c(100,10,10)) #implicit vectorization
add(1, 1) #Fails due to missing argument
```
AddNEW now add(x, y = 1):
```{R}
addNew <- function(x ,y = 1){
  x + y
}
```

When the y value is not defined it assumes the detailt valuye of 1, but when we do provide it a value, then it overrides the default value within the argument.
```{R}
addNew(100) #returns 101
```

## Second Function
Let's try something interesting: Make a sequence generating tool..

The `sample()` function is an inbuilt function in R, that has 2 arguments and randomly samples some input. X is the sample space, size is the sample size, and replace (T/F) is used to derive repeats.

```{R}
sample(1:10, size = 4)

```

>Q. Generate 9 random numbers from the input vector 1:10: Cannot do 12, as the sample size cannot be larger than the population.

```{R}
sample(1:10, size = 9)

```

>Q. Generate 12 random numbers from the input vector 1:10: 
Cannot do 12, as the sample size cannot be larger than the population.

```{R}
sample(1:10, size=12, replace = TRUE)

```





>Q Write a code for `Sample()` function that generates nucleotide sequences of length 6:

```{R}
sample(c('A','C','G','T'), size = 6, replace = TRUE)
```


>Q Write a first function `generate_dna()` that returns the user a specified length DNA sequence:

```{R}
generate_dna <- function(len){
  sample(c('A','C','G','T'), size = len, replace = TRUE)
}

```

```{R}
generate_dna(100)
```


> **Key-Points**
Each function in R looks fundamentally the same it has the following structure:

```
name <- function(args){
  function body
}
```

>Functions can have multiple inputs, some of which are **required** and some **optional**. With optional arguments having a set default value.



> Q. Modify and prove our `generate_dna()` function to return it's generated sequence in a more standard form: like "AGTAGTA" rather than vector

```{R}
generate_dna <- function(len, fasta = TRUE){
  ans <- sample(c('A','C','G','T'), size = len, replace = TRUE)
  
  if (fasta){
    cat("Single-element vector output")
    ans <- paste(ans, collapse = "")
  }else{
    cat("Multi-element vector output")
  }
  
  return (ans)
}

generate_dna(10)
generate_dna(10, FALSE)

```



The `paste()` function is used to join up or stick together elements in a vectorized structure.

```{R}
paste(c("Sathvik", "Barry"), "loves R", sep = "")
```


Flow control is where the R brain goes into your code: Conditionals!
```{R}
good_mood <- FALSE

if (good_mood){
  cat("Great!")
} else{
  cat("Bummer!")
}

```

## A Protein generating function:

>Q. Write a function that generates a user- specified length protein sequence.

```{R}
generate_protein <- function(len){
  ans <- sample(c("A","C","D","E","F","G","H","I","K","L","M","N","P","Q","R","S","T","V","W","Y"),
         size = len, replace = TRUE)
  return(cat(ans, sep= ""))
}

generate_protein(14)

```

>Q. Use that function to generate a random protein sequence between length 6 and 12.

```{R}
random_protein <- function(){
  randomN = sample (6:12, size = 1)
  sample(c("A","C","D","E","F","G","H","I","K","L","M","N","P","Q","R","S","T","V","W","Y"),
         size = randomN, replace = TRUE)
}

random_protein()
```

```{R}
generate_protein(6)
generate_protein(7)
generate_protein(8)
generate_protein(9)
generate_protein(10)
generate_protein(11)
generate_protein(12)
```
>Q. Are any of your sequences unique i.e not found anywhere in nature?

**YES**, when I ran BLASTP on (Len = 12)FKMRQHDWMISS, there's lower query cover; therefore they're unique. Though (6-9 are not unique.)

```{R}
for (i in 6:12){
  cat(i, "\n")
  cat(generate_protein(i), "\n")
}

```

